$\epsilon$-explicit analysis of DPG for convection-dominated diffusion

### 2012-04-10 docs/NewBCRobustness_paper/paper.tex
removed a typo (piecewise constant, when I just meant constant)
git-svn-id: https://svn.ices.utexas.edu/repos/camellia@460 36274a68-8941-4c93-80ef-affaed5139da

Consider now the 1D analogue of the estimate given by Lemma~\ref{lemma_stream}, which is necessary for robust control over the error in $\|u\|_{L^2}$.  In 1D, $\|\beta\cdot\grad v\| \lesssim \|g\|$ reduces to the inequality 
\[
\|v'\|\lesssim \|g\|, \quad g\in L^2{\Omega_h}
\]
The adjoint problem corresponding to Lemma~\ref{lemma_stream} in section~\ref{sec:strategy3} is likewise reduced in 1D to the scalar equation
\begin{align}
\epsilon v'' + v' &= -g \label{adjoint_1D}
\end{align}
with $v\in H^1_0\left((0,1)\right)$.  Multiplying this equation by $v'$ and integrating over $\Omega_h$ gives
\[
\int_0^1 \epsilon v''v' + \|v'\|_{L^2}^2 = -\int_0^1 g v'
\]
Integrating by parts and Young's inequality give
\begin{align*}
-\left.\frac{\epsilon}{2} v'^2\right|_0^1 + \|v'\|_{L^2}^2 &\leq \frac{1}{2}\|g\| + \frac{1}{2}\|v'\|
\end{align*}
implying that
\begin{align*}
\frac{\epsilon}{2} v'(0)^2 + \|v'\|_{L^2}^2 & \lesssim \|g\| 
\end{align*}
Let us restrict ourselves to the cases where $v$ is sufficiently smooth for $v'(0)$ to be well defined.  Taking $g=1$ (corresponding to a piecewise constant approximation) we can solve \eqref{adjoint_1D} exactly to see that 
\[
v(x) = \frac{e^{-\frac{x}{\epsilon}}}{e^{\frac{1}{\epsilon}}-1}\left(e^{\frac{1}{\epsilon}}\left(e^{\frac{x}{\epsilon}}-1\right) + \left(e^{\frac{1}{\epsilon}}-1\right)e^{\frac{x}{\epsilon}}x\right)
\]
Plotting the solution shows that $v$ develops strong boundary layers of width $\epsilon$ near the inflow. Consequently, $\frac{\epsilon}{2} v'(0)^2 \approx \epsilon^{-1}$, indicating already a lack of robustness in the bound of $\|v'\|$ by $\|g\|$ for constant data\footnote{Unlike the case of Dirichlet boundary conditions, the inflow condition on $ \widehat{f} = u(0)-\epsilon u'(0)$ induces an adjoint boundary condition $\tau(0)=0$, or equivalently $v'(0) = 0$, removing the non-robust term from the estimate.}. The robust bound from below of $g$ constant corresponds to the robust control of the approximation of field variable $u$ by piecewise constant trial functions. More detailed 1D error bounds for Dirichlet boundary conditions are provided in \cite{DPG3}, and also indicate a lack of robustness when using the standard broken Sobolev norm as the test norm on $\tau$ and $v$.

>>>>

Consider now the 1D analogue of the estimate given by Lemma~\ref{lemma_stream}, which is necessary for robust control over the error in $\|u\|_{L^2}$.  In 1D, $\|\beta\cdot\grad v\| \lesssim \|g\|$ reduces to the inequality 
\[
\|v'\|\lesssim \|g\|, \quad g\in L^2{\Omega_h}
\]
The adjoint problem corresponding to Lemma~\ref{lemma_stream} in section~\ref{sec:strategy3} is likewise reduced in 1D to the scalar equation
\begin{align}
\epsilon v'' + v' &= -g \label{adjoint_1D}
\end{align}
with $v\in H^1_0\left((0,1)\right)$.  Multiplying this equation by $v'$ and integrating over $\Omega_h$ gives
\[
\int_0^1 \epsilon v''v' + \|v'\|_{L^2}^2 = -\int_0^1 g v'
\]
Integrating by parts and Young's inequality give
\begin{align*}
-\left.\frac{\epsilon}{2} v'^2\right|_0^1 + \|v'\|_{L^2}^2 &\leq \frac{1}{2}\|g\| + \frac{1}{2}\|v'\|
\end{align*}
implying that
\begin{align*}
\frac{\epsilon}{2} v'(0)^2 + \|v'\|_{L^2}^2 & \lesssim \|g\| 
\end{align*}
Let us restrict ourselves to the cases where $v$ is sufficiently smooth for $v'(0)$ to be well defined.  Taking $g=1$ (corresponding to a constant approximation) we can solve \eqref{adjoint_1D} exactly to see that 
\[
v(x) = \frac{e^{-\frac{x}{\epsilon}}}{e^{\frac{1}{\epsilon}}-1}\left(e^{\frac{1}{\epsilon}}\left(e^{\frac{x}{\epsilon}}-1\right) + \left(e^{\frac{1}{\epsilon}}-1\right)e^{\frac{x}{\epsilon}}x\right)
\]
Plotting the solution shows that $v$ develops strong boundary layers of width $\epsilon$ near the inflow. Consequently, $\frac{\epsilon}{2} v'(0)^2 \approx \epsilon^{-1}$, indicating already a lack of robustness in the bound of $\|v'\|$ by $\|g\|$ for constant data\footnote{Unlike the case of Dirichlet boundary conditions, the inflow condition on $ \widehat{f} = u(0)-\epsilon u'(0)$ induces an adjoint boundary condition $\tau(0)=0$, or equivalently $v'(0) = 0$, removing the non-robust term from the estimate.}. The robust bound from below of $g$ constant corresponds to the robust control of the approximation of field variable $u$ by a constant. More detailed 1D error bounds for Dirichlet boundary conditions are provided in \cite{DPG3}, and also indicate a lack of robustness when using the standard broken Sobolev norm as the test norm on $\tau$ and $v$.
